\vspace*{\fill}

\noindent \textbf{\Huge Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

At this point I would like to thank all those people who have
supported me professionally, technically and morally throughout the
development of this research.  While I cannot list them all here,
certain groups and individuals are deserving of special
acknowledgement.

This work was carried out in the Fusion Technology Institute [FTI] of
the University of Wisconsin-Madison's Engineering Physics [EP]
Department, with partial support from the United States Department of
Energy.  I gratefully acknowledge the support and feedback of all my
FTI and EP colleagues.

Despite the short opportunity I had to know him, the late Prof.
Emeritus Charles Maynard provided much inspiration for this work.  He
always insisted that loops were usually not important, and when they
were, the solution could be easily found.  I am happy to have proven
him correct.

Special thanks are due to Professor Douglass Henderson for presiding
over this research.  We have both learned much more about the nature
of this problem through many hours of debate and discussion over the
last six years.  \ALARA\ would never have been completed without his
guidance.

In addition, Jim Sisolak, Jeff Crowell and Prof. Jake Blanchard have
all, at times, been good listeners to help me focus my ideas and
develop my methods.

Finally, a very special thanks and recognition for my wife, Laurie
Nagus for putting up with the long hours, for listening to my
complaining, and most of all, for always believing that I could do it.






